---
title: "Fish Recruitment Analysis"
author: "Kelsey Webber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    css: resources/AIMS-style.css
    toc: true
    toc-float: true
    number-sections: true
    number-depth: 3
    embed-resources: true
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
  tbl-title: '**Table**'
  tbl-labels: arabic
engine: knitr
bibliography: resources/references.bib
output_dir: "docs"
---

```{r chunks, results='markdown', eval=TRUE, echo = FALSE}
knitr::read_chunk('../R/functions.R')
knitr::read_chunk('../R/analysis_recruitment.R')

```

# Preparations

- Load any helper functions (functions.R)

```{r loadFunctions, results='markdown', eval=FALSE}
```

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
```


- Load required packages (functions.R)

```{r loadPackages, results='markdown', eval = TRUE, warnings = FALSE, message = FALSE}
```

- Prepare paths (functions.R)

```{r preparePaths, results='markdown', eval = TRUE}

```

# Fish Recruitment

## Load data

1. Load the data
```{r FishRecruitment readData, results = 'markdown', eval = TRUE}
```

2. Glimpse the data
```{r recruitment glimpse, results='markdown', eval=TRUE}
```

```{r recruitment report table, results='markdown', eval=TRUE}
```

```{r factorise and clean, results = 'markdown', eval = TRUE}
```

## EDA

1. Algae EDA

```{r recruitment algae EDA, results = 'markdown', eval = TRUE}
```
A bit of an outlier in Length - One plot (W5) had an extra long 52cm thallus. I checked in the video and this was not a typo. However its weight was not super high (70g). Lengths otherwise fairly similar, around 30cm

Weights per plot: BQ < DL <= BH <= DM < W

2. Fish EDA
```{r recruitment fish EDA, results = 'markdown', eval = TRUE}
```