---
title: "Fish Recruitment Analysis"
author: "Kelsey Webber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    css: resources/AIMS-style.css
    toc: true
    toc-float: true
    number-sections: true
    number-depth: 5
    embed-resources: true
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
  tbl-title: '**Table**'
  tbl-labels: arabic
engine: knitr
bibliography: resources/references.bib
output_dir: "docs"
---

```{r chunks, results='markdown', eval=TRUE, echo = FALSE}
knitr::read_chunk('../R/functions.R')
knitr::read_chunk('../R/analysis_recruitment.R')

```

# Preparations

- Load any helper functions (functions.R)

```{r loadFunctions, results='markdown', eval=FALSE}
```

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
```


- Load required packages (functions.R)

```{r loadPackages, results='markdown', eval = TRUE, warnings = FALSE, message = FALSE}
```

- Prepare paths (functions.R)

```{r preparePaths, results='markdown', eval = TRUE}

```

# Load Fish Recruitment data

1. Load the data
```{r FishRecruitment readData, results = 'markdown', eval = TRUE}
```

2. Glimpse the data
```{r recruitment glimpse, results='markdown', eval=TRUE}
```

```{r recruitment report table, results='markdown', eval=TRUE}
```

3. Factorise and clean
```{r factorise and clean, results = 'markdown', eval = TRUE}
```

# EDA

## Algae EDA

```{r recruitment algae summary table, results = 'markdown', eval = TRUE}
```

```{r recruitment algae summary figures, results = 'markdown', eval = TRUE}
```
A bit of an outlier in Length - One plot (W5) had an extra long 52cm thallus. I checked in the video and this was not a typo. However its weight was not super high (70g). Lengths otherwise fairly similar, around 30cm

Weights per plot: BQ < DL <= BH <= DM < W

## Fish EDA

### 1 Total abundance
```{r fish EDA total abundance, results = 'markdown', eval = TRUE}
```
Total abundance seems to follow a similar pattern to the biomass of macroalgae. It’s also very nice that the variance doesn’t seem to differ among treatments

1.1 Total abundance over time
```{r fish EDA abundance time, results = 'markdown', eval = TRUE}
```
Does abundance/diversity/biomass (size) change differently over time among treatments?

Yes (Abundance) – W treatment seems to continuously increase, while other treatments (esp DL and BQ) flatten early. Carrying capacity being reached earlier? DM shows a hump in abundance

1.2 Abundance vs Algal biomass
```{r fish EDA abundance vs biomass, results = 'markdown', eval = TRUE}
```
Could the plot biomass be used as a predictor of total abundance?- yeah seems so

### 2 Species Richness

```{r fish EDA Species richness, results = 'markdown', eval = TRUE}
```
Similarly to abundance, richness seems to follow a similar pattern to the biomass of macroalgae.


2.1 Species richness over time
```{r fish EDA richness vs time, results = 'markdown', eval = TRUE}
```
Does abundance/diversity/biomass (size) change differently over time among treatments?

For diversity,  W again seems to continuously increase, while other treatments (esp DL and BQ) flatten early. DM shows an even more prominent hump in abundance around day 8.

2.2 Species Richness vs Algal biomass
```{r fish EDA richness vs biomass, results = 'markdown', eval = TRUE}
```

Could the plot biomass be used as a predictor of diversity ?- yeah seems so

### 3 Find Common Species
```{r fish EDA commonness, results = 'markdown', eval = TRUE}
```
The top 6 (Halichoeres miniatus to Siganus fuscescens) collectively accounted for 90% of total abundance, each individually accounted for more than 2% of total abundance and appeared every day. Only Steothjulis strigiventer also appeared every day (but why analyse it if not something else more abundant?) Using just the top 6 or less makes sense

### 4 Common species patterns
```{r fish EDA common species1, results = 'markdown', eval = TRUE}
```

Petroscirtes and Siganus doliatus and maybe Pomacentrus tripunctatus show a somewhat similar patttern to that of plot biomass, but lots of skew. Halichoeres miniatus shows a fairly similar abundance among treatments (although I wonder if the size distributions may vary?) Maybe not enough data for Lethrinus atkinsoni and Siganus fuscescens

4.1 Common species over time
```{r fish EDA common species2, results = 'markdown', eval = TRUE}
```

H. miniatus, Petroscirtes sp. and tripunctatus accumulated over time, particularly in treatments W, BH and DM (H. miniatus), W (Petroscirtes) and all except BQ in P. tripunctatus. S doliatus declined in abundance over time to making W, BH and DM, which started off with more, approximately similar to the other treatments after ~6 days. L atkinsoni and S. fuscescens show no clear pattern (not enough data)

4.2 Common sp vs algal biomass
```{r fish EDA common species3, results = 'markdown', eval = TRUE}
```
Unlikely to be worth exploring L. atkinsoni and S. fuscescens, but the rest show a bit of a positive relationship (including H. miniatus kind of). Petroscirtes sp. the strongest relationship

### 5 Size distributions

```{r fish EDA sizes1, results = 'markdown', eval = TRUE}
```
This plot shows the mean number of fish of each length category (bin width 1) per day for each treatment, all fish combined.
While the overall centre of mass is similar among treatments (around 2-5), the height of the peaks appear to increase with increasing plot biomass (treatments BQ=DL<BH<DM<W)

```{r fish EDA sizes2 common species, results = 'markdown', eval = TRUE}
```
These plots shows the mean number of fish of each length category (bin width 1) averaged by day and treatment, separated by species for the 6 most common.
Halichoeres miniatus shows a relatively similar distribution across treatments. However, Petroscirtes sp., Siganus doliatus and Pomacentrus tripunctatus show similar patterns to the overall size distributions - greater mass around ~2.5cm as plot biomass of sargassum increases (treatments BQ=DL<BH<DM<W)

It probably isn't worth exploring S.fuscescens and L. atkinsoni further (too many 1s and not much else)


```{r fish EDA sizes2 over time, results = "markdown", eval = TRUE, quiet = TRUE}
```
These plots show the mean number of fish of each length category averaged by Treatment and plotted for each day separately
Not much of a pattern really to see - no obvious shift in size distributions over time for all fish combined, Petroscrites sp. or for Siganus doliatus


# Univariate modelling

## Total abundance
```{r recruitment univariate setup data, results = 'markdown', eval = TRUE}
```


### Fit model
```{r recruitment univariate abundance initial models, results = 'markdown', eval = TRUE}
```
The 'Treatment' model with performed best. Plot weight model and all models with density included performed poorly. Note it wouldn't make sense to include density and treatment, but plot weight did vary within treatments. However the models including Treatment and plot.weight performed poorly 

### Validate/Refit
```{r recruitment univariate abundance validate, results = 'markdown', eval = TRUE}
```
residuals look quite  good.
However, there appears to be definite autocorrelation
No zero inflation


```{r recruitment univariate abundance refit autocorrelation, results = 'markdown', eval = TRUE}
```
autocorrelation model fit straight away, and is the definite winner by AICc despite higher df

```{r recruitment univariate abundance revalidate, results = 'markdown', eval = TRUE}

```
The autocorrelation model's residuals didn't meet the assumptions, particularly uniformity. It also still shows autocorrelation. The result of the Dispersion test indicates UNDER-dispersion - of less concern than overdispersion, as rather than inflating the type I error rate, it shrinks it - more conservative

However: (from https://rdrr.io/cran/DHARMa/man/testTemporalAutocorrelation.html)

Note
Standard DHARMa simulations from models with (temporal / spatial / phylogenetic) conditional autoregressive terms will still have the respective temporal / spatial / phylogenetic correlation in the DHARMa residuals, unless the package you are using is modelling the autoregressive terms as explicit REs and is able to simulate conditional on the fitted REs. This has two consequences

If you check the residuals for such a model, they will still show significant autocorrelation, even if the model fully accounts for this structure.

Because the DHARMa residuals for such a model are not statistically independent any more, other tests (e.g. dispersion, uniformity) may have inflated type I error, i.e. you will have a higher likelihood of spurious residual problems.

There are three (non-exclusive) routes to address these issues when working with spatial / temporal / other autoregressive models:

Simulate conditional on the fitted CAR structures (see conditional simulations in the help of simulateResiduals)

Rotate simulations prior to residual calculations (see parameter rotation in simulateResiduals)

Use custom tests / plots that explicitly compare the correlation structure in the simulated data to the correlation structure in the observed data.

### Partial Plot
```{r recruitment univariate abundance partial, results='markdown', eval = TRUE}

```


### Bayesian Model
#### 1 Initial priors
```{r recruitment univariate abundance priors1, results = 'markdown', eval = TRUE}

```
The default priors were mostly flat for the effects and student_t for the sds. A prior I haven't used before appeared - class 'ar' - flat, with an upper bound of 1 and a lower bound of -1.
From the plots and eda, above, an appropriate estimate of the abundance in group BH (the default first) would be 8. On the log scale, log 8 = `r log(8)` ~ 2.1 (the same as brms would give for the median of the student_t for the Intercept). However the dispersion of 2.5 is likely too wide.
In addition to an sd overall, an sd for the intercept and an sd by group = plotID, an additional 'sderr' term exists. This is the "standard deviation of latent residuals in latent ARMA models".

for the intercept, a more appropriate prior might be N(2,1)

For the slope, sd(RESPONSE/sd(PREDICTOR) – gives a very crude approximation of the sd of the slope of that relationship. In this case our predictor is Categorical, so we have to use dummy code (0s and 1s)
```{r recruitment univariate abundance priors2, results = 'markdown', eval = TRUE}

```

The sds are the same for each treatment because the study was balanced.

For the variance argument for the slope prior, a bit more than 4, say 4.5 or 5 (to allow a bit more width) would probably be appropriate
i.e. N(0, 5)

For the sd prior, a cauchy (0,2) will be used to begin with. I'll do the same with "sderr"

#### 2 Fitting
2.1 Priors only
```{r recruitment univariate abundance fit brm1.prior, results = 'markdown', eval = FALSE}

```
The above code model did fit, but I had trouble with using ggpredict on it (see below) 

```{r recruitment univariate abundance brm1 error and fix attempt, results = 'markdown', error = TRUE}

```
Changing the class of 'Date' and refitting the model did not fix the issue

2.2 Fit alternative model (no ar term)
```{r recruitment univariate abundance fit brm2, results = 'markdown', eval = FALSE}

```

The model fit, however and ggpredict worked for this model and allowed refinement of the priors for the b's
```{r recruitment univariate abundance brm2 ggpredict, results = 'markdown', eval = TRUE}
```
a normal(0,2) was deemed probably more appropriate for b


2.3 Fit model, sampling from data and from priors
```{r recruitment univariate abundance fit brm1a_b, results = 'markdown', eval = FALSE}
```
The eventual priors and model "abnd.brm1b" was chosen because: 
ar: there was no reason to expect a negative autocorrelation could occur, and reducing the prior slightly improved std.errors and ess's
sd and sderr: reducing the priors reduced the magnitude of difference between the posterior and prior. minimal effect on results, possibly more ess and smaller CI



```{r recruitment univariate abundance vars, results = 'markdown', eval = TRUE, error = TRUE, cache = TRUE}

```

The model fit, ggpredict had trouble, and there seems to be an error estimate for every data point (err[1]...err[450]). Compare with:

`r abnd.brm2.prior %>% get_variables()` 

(the variables of the model without the ar term)

#### 3 Prior Checks

```{r recruitment univariate abundance brm1b prior checks, results = 'markdown', eval = TRUE, cache = TRUE}

```
Prior for the intercept and treatment seem appropriate. sd and sderr could potentially shrink if further diagnostics reveal problems

#### 4 MCMC diagnostics
```{r recruitment univariate abundance brm1b MCMC diagnostics, results = 'markdown', eval = TRUE}
```
Chains are well mixed, autocorrelation factors are low, rhats are around 1 and less than 1.05, effective sample sizes for the posteriors are all close to 1 and well above 0.5 and posterior distributions of the separate chains resemble one another

"All chains were well mixed and converged (all Rhat values <1.05)"

#### 5 DHARMA residuals

```{r recruitment univariate abundance DHARMA residuals, results = 'markdown', eval = TRUE}
```
The residuals don't look too good - similarly to the Frequentist approach, both dispersion (UNDERdispersion) and uniformity appear to be violated

However, this again is likely caused by not simulating Conditional residuals


### Model Investigation
#### 1 Frequentist models
```{r recruitment univariate abundance summary, results = 'markdown', eval = TRUE, error =TRUE}

```
Clear differences among treatments. 

Estimate of BH mean abundance: `r exp(2.1)`. Estimate for W is `r exp(.56)` x higher (`r exp(2.1+.56)`)

R-square estimates indicate the model without the random component of PlotID explains approximately 50% of the variance. With the random effect it explains 62% (using the trigamma method as this model has a log link, package authors recommend this). R-square calculation didn't work for the ac model using this particular method

Planned contrasts: from Murray Logan Course notes "In frequentist analyses you can make nlevels - 1 contrasts without p adjustment - as long as they're independent." - however I've read one other source suggesting one should still apply a correction even then. I'll leave this for now. If running Bayesian, I can make as many contrasts as I want

#### 2 Bayesian Model
2.1 Summary
```{r recruitment univariate abundance brm summary, results = 'markdown', eval = TRUE, error =TRUE}

```
although 'ggpredict()' is having some trouble, ggemmeans works well

The summary stat table shows differences amongst treatments, will explore further contrasts (e.g. W vs the rest) next because I can

the estimate of ar(1) is slightly lower compared to the Frequentist analysis

```{r recruitment univariate abundance brm r square, results = 'markdown', eval = TRUE}

```


R-squared - 95% confidence the amount of variance explained falls between 58.6 and 68.6% (median 63%)
with the random effects set to 0, the model explains about 56% (conditioned on an 'average' random effects model, i.e. conditioned on 0, or marginal in a linear model)

2.2 Hypothesis testing

```{r recruitment univariate abundance all contrasts, results = 'markdown', eval = TRUE, warning=TRUE}

```
There is very strong evidence (>99%) that treatment BH had higher fish recruitment than BQ and DL, that W had higher recruitment than BH, BQ and DL, and that DM had higher recruitment than BQ and DL. There was strong evidence (>95%) that W had higher recruitment than DM.There was evidence (>90%) that DM had higher recruitment than BH. There was no evidence (<90%) for a difference between BQ and DL

Compare:
Natural vs Reduced Biomass: Mean of DL, DM and W vs Mean of BH and BQ
High vs Reduced Density: Mean of BH, BQ and W vs Mean of DL and DM
All vs Low Density: DL vs the rest
```{r recruitment univariate abundance planned contrasts, results = 'markdown', eval = TRUE, warning=TRUE}

```
There was strong evidence that reducing Sargassum biomass, regardless of Sargassum density, resulted in fewer fish recruits. There was no evidence that reducing density, below 9 per plot, regardless of thallus biomass, reduced fish recruitment.

However there was strong evidence that total recruitment was reduced at the low extremes of both biomas and density compared to the average recruitment of the other treatments. (perhaps not the most sensible contrast)

```{r recruitment univariate abundance summary figure, results = 'markdown', eval = TRUE}

```

See bglm_example5.html for fancy figure code. Need to fix X axis labels in g2 and perhaps make the 'stat_slab' object end at the 95% intervals in g1


## Species Richness

### Fit model
```{r recruitment univariate sp fit, results = 'markdown', eval = TRUE}
```
As with abundance, the Treatment + RE of plot model performed best, despite consumingn more df

###Validate/Refit
```{r recruitment univariate sp validate, results = 'markdown', eval = TRUE}

```
The residuals are showing signs of underdispersion (i.e. results more conservative). The Levene test is also warning that the homogeneity of variance assumption is violated. However the box plots look ok to me


```{r recruitment univariate sp refit autocor, results = 'markdown', eval = TRUE}

```
Despite there being evidence of autocorrelation, the ac model performed about the same (within 2 AICc) as the previous model. However I think the autocorrelation is important enough to need dealing with

```{r recruitment univariate sp revalidate, results = 'markdown', eval = TRUE}

```
Similar problems evident (undispersion), but without the red flag for the Levene test


### Partial Plot
```{r recruitment univariate sp partial, results='markdown', eval = TRUE}

```

### Bayesian Model
####1 Initial priors
```{r recruitment univariate sp priors1, results = 'markdown', eval = TRUE}

```
As the model formula is so similar to that of the abundance model, the same priors will need specification.

For the Intercept, an appropriate prior might be 1, 0.4

Again, the standard deviation of the response divided by the standard deviation of the predictor will be used to estimate an appropriate prior for the slopes:
```{r recruitment univariate sp priors2, results = 'markdown', eval = TRUE}

```
N(0,2) might be appropriate for the slopes

Again, I'll try a cauchy(0,2 for sd and sderr


```{r recruitment univariate sp prior fit, results = 'markdown', eval = FALSE}
```

```{r recruitment univariate sp brm1.prior check, results = 'markdown', eval = TRUE}
```

Priors quite wide compared with data. Remember this is not in itself a problem, unless we find evidence later

####2 Fitting

```{r recruitment univariate sp brmsfit, results = 'markdown', eval = FALSE}
```
2048 divergent transitions occurred, out of a possible `r 4000/5 * 3`. Also note that the size of of both brmsfits (sp.brm1 and sp.brm1a) are apparently of size 0B in the environment ( though their saved file sizes seem similar to those of the abnd models)


####3 Prior Checks
```{r recruitment univariate sp brm1a prior checks, results = 'markdown', eval = TRUE}

```
Prior seems appropriate for effects, and for the intercept they could maybe go slightly wider. sd and sderr priors are very wide, so could come in a bit (but see prior checks for abundance - very similar). Prior for ar however looks identical to the posterior, despite autocorrelation being similar or greater (0.98 according to  summary(sp.glmmTMB.ac), Durbin-Watson test p < 2.2e-16) than for abundance


####4 MCMC Diagnostics

```{r recruitment univariate sp MCMC1, results = 'markdown', eval = TRUE}

```
Trace plots look good, resembling noise and each chain resembles the rest
Rhats fine (close to 1, not over 1.05)
Density overlay looks pretty good


```{r recruitment univariate sp MCMC2, results = 'markdown', eval = TRUE}

```

AC - all have values of >0.2 at first lag - increase thinning?
ESS - most values below 0.5, none over 0.6 - priors too wide? adapt delta adjustment?
Density plot shows no real mode for ar

Options:
adapt_delta to 0.99
Shrink priors for sd and sderr
ar prior??
adjust thinning factor

####4.1 Refitting, rechecking
```{r recruitment univariate sp brms refit, results = 'markdown', eval = FALSE}

```
2121 divergent transitions

```{r recruitment univariate sp recheck priors and mcmc, results = 'markdown', eval = TRUE, cache = TRUE}

```
ess and ac now acceptable

```{r recruitment univariate sp recheck recheck mcmc2, results = 'markdown', eval = TRUE}

```
all other mcmc diagnostics now acceptable, but the posteriors for ac are worrying - no consistent mode (particularly not near the frequentist derived estimate of 0.98)

####5 DHARMA residuals

```{r recruitment univariate sp DHARMA residuals, results = 'markdown', eval = TRUE}

```
Showing underdispersion again, but this can be explained as for abundance

### Model investigation
####1 Frequentist

```{r recruitment univariate sp frequentist summary, results = 'markdown', eval = TRUE, error = TRUE}
```
Significant differences between intercept (BH) and DL, W. 

0.98 ar1

The model without the ar term is explaining 36% of the variance in species richness with the RE, 26% without

####2 Bayesian
```{r recruitment univariate sp bayesian summary, results = 'markdown', eval = TRUE, error = TRUE}
```
I'm concerned about the estite for ar - it's basically 0.5, wit very wide confidence intervals (almost to the width of the prior uniform(0,1) dist), not anywhere close to the frequentist estimate