---
title: "Fish Recruitment Analysis"
author: "Kelsey Webber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    css: resources/AIMS-style.css
    toc: true
    toc-float: true
    number-sections: true
    number-depth: 4
    embed-resources: true
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
  tbl-title: '**Table**'
  tbl-labels: arabic
engine: knitr
bibliography: resources/references.bib
output_dir: "docs"
---

```{r chunks, results='markdown', eval=TRUE, echo = FALSE}
knitr::read_chunk('../R/functions.R')
knitr::read_chunk('../R/analysis_recruitment.R')

```

# Preparations

- Load any helper functions (functions.R)

```{r loadFunctions, results='markdown', eval=FALSE}
```

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
```


- Load required packages (functions.R)

```{r loadPackages, results='markdown', eval = TRUE, warnings = FALSE, message = FALSE}
```

- Prepare paths (functions.R)

```{r preparePaths, results='markdown', eval = TRUE}

```

# Load Fish Recruitment data

1. Load the data
```{r FishRecruitment readData, results = 'markdown', eval = TRUE}
```

2. Glimpse the data
```{r recruitment glimpse, results='markdown', eval=TRUE}
```

```{r recruitment report table, results='markdown', eval=TRUE}
```

3. Factorise and clean
```{r factorise and clean, results = 'markdown', eval = TRUE}
```

# EDA

## Algae EDA

```{r recruitment algae summary table, results = 'markdown', eval = TRUE}
```

```{r recruitment algae summary figures, results = 'markdown', eval = TRUE}
```
A bit of an outlier in Length - One plot (W5) had an extra long 52cm thallus. I checked in the video and this was not a typo. However its weight was not super high (70g). Lengths otherwise fairly similar, around 30cm

Weights per plot: BQ < DL <= BH <= DM < W

## Fish EDA

### 1.0 Total abundance
```{r fish EDA total abundance, results = 'markdown', eval = TRUE}
```
Total abundance seems to follow a similar pattern to the biomass of macroalgae. It’s also very nice that the variance doesn’t seem to differ among treatments

1.1 Total abundance over time
```{r fish EDA abundance time, results = 'markdown', eval = TRUE}
```
Does abundance/diversity/biomass (size) change differently over time among treatments?

Yes (Abundance) – W treatment seems to continuously increase, while other treatments (esp DL and BQ) flatten early. Carrying capacity being reached earlier? DM shows a hump in abundance

1.2 Abundance vs Algal biomass
```{r fish EDA abundance vs biomass, results = 'markdown', eval = TRUE}
```
Could the plot biomass be used as a predictor of total abundance?- yeah seems so

### 2.0 Species Richness

```{r fish EDA Species richness, results = 'markdown', eval = TRUE}
```
Similarly to abundance, richness seems to follow a similar pattern to the biomass of macroalgae.


2.1 Species richness over time
```{r fish EDA richness vs time, results = 'markdown', eval = TRUE}
```
Does abundance/diversity/biomass (size) change differently over time among treatments?

For diversity,  W again seems to continuously increase, while other treatments (esp DL and BQ) flatten early. DM shows an even more prominent hump in abundance around day 8.

2.2 Species Richness vs Algal biomass
```{r fish EDA richness vs biomass, results = 'markdown', eval = TRUE}
```

Could the plot biomass be used as a predictor of diversity ?- yeah seems so

### 3 Find Common Species
```{r fish EDA commonness, results = 'markdown', eval = TRUE}
```
The top 6 (Halichoeres miniatus to Siganus fuscescens) collectively accounted for 90% of total abundance, each individually accounted for more than 2% of total abundance and appeared every day. Only Steothjulis strigiventer also appeared every day (but why analyse it if not something else more abundant?) Using just the top 6 or less makes sense

### Common species patterns
