---
title: "fish.recruitment.multivariate"
author: "Kelsey Webber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    css: ../docs/resources/AIMS-style.css
    toc: true
    toc-float: true
    number-sections: true
    number-depth: 5
    embed-resources: true
    code-fold: false
    code-tools: true
    code-summary: "Show the code"
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
  tbl-title: '**Table**'
  tbl-labels: arabic
engine: knitr
output_dir: ../docs
---

```{r chunks, results='markdown', eval=TRUE, echo = FALSE, cache = TRUE}
knitr::read_chunk('../R/functions.R')
knitr::read_chunk('../R/analysis_recruitment.R')

```

# Preparations

- Load any helper functions (functions.R)

```{r loadFunctions, results='markdown', eval=FALSE}
```

```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
```


- Load required packages (functions.R)

```{r loadPackages, results='markdown', eval = TRUE, warnings = FALSE, message = FALSE}
```

- Prepare paths (functions.R)

```{r preparePaths, results='markdown', eval = TRUE}

```

# Load Fish Recruitment data

1. Load the data
```{r recruitment multivariate readData, results = 'markdown', eval = TRUE}
```


2. Set up wide data
```{r ecruitment multivariate data set up, results = 'markdown', eval = TRUE}
```


# MDS

```{r recruitment multivariate end mds, results = 'markdown', eval = TRUE}

```
on distance: Stress is at an acceptable level (0.14). 2 convergent solutions found


#Adonis
```{r recruitment multivariate end adonis, results = 'markdown', eval = TRUE}

```
significant adonis result, but this could be driven by differences in dispersion
BH different to W and DL

# EnvFit - Species vectors
```{r recruitment multivariate end envfit, results = 'markdown', eval = TRUE}

```
Petroscirtes, P.T, SD, LA, Steth strig and steth interrupta and SF significant. Lethrinus variegatus p = 0.057

# Plot MDS
```{r recruitment multivariate end mds ggplot, results = 'markdown', eval = TRUE}

```


```{r recruitment multivariate end disper, results = 'markdown', eval = TRUE}

```
Dispersion among treatments look ok?? or W may be different to DL- permuted p value of ~0.01 for their contrast
